Physiological plasticity and variance in physiological rates vary across ectotherm populations

### 2024-07-13 ms.qmd
huge overhaul and re-organisation of code. bring lnVR to main MS. Re-rendered after a few failures as there were code bugs. Need to fix results sections now, mainly re-write and add a fe wmore overall comparisons across habiats

@@

@@

@@

@@

@@

@@

@@

####################
# lnCVR CV
####################
dat_textcvr <- data.frame(label = c(text(slope_f_4.9), 
                                 text(slope_m_4.9), 
                                 text(slope_t_4.9)),
                      label2 = c(text(slope_f_4.9, b = FALSE), 
                                 text(slope_m_4.9,b = FALSE), 
                                 text(slope_t_4.9,b = FALSE)),
                           p = c(p_text(slope_f_4.9), 
                                 p_text(slope_m_4.9), 
                                 p_text(slope_t_4.9)),
                       title = c("Freshwater", "Marine", "Terrestrial"), 
                      condition = c("f", "m", "t"),
                      x = rep(0, 3),
                      y = rep(-4, 3))

# Freshwater
  yadj2 = -6
  x_f = 0.6
  p_f_cv <-  bubble_plot(model4.9_f, mod = "cv_c", group = "spp", ylab = TeX("$lnCVR_{Q_{10_{acclim}}}$"), xlab = "Thermal Coefficient of Variation (CV)", g =FALSE, k = FALSE) + 
      geom_point(data = lnCVRQ10_data_wild_acclim_f, aes(x = cv_c, y = effect_size, size = 1/sqrt(V)), fill = col[1], colour = "black", stroke = 0.2, shape = 21) + 
    annotate("text", x = dat_textcvr$x[1]+x_f, y = yadj2, label = TeX(dat_textcvr$p[1]), size = 8) +
    annotate("text", x = dat_textcvr$x[1]+x_f, y = yadj2+0.8, label = TeX(dat_textcvr$label2[1]), size = 8) +
    annotate("text", x = dat_textcvr$x[1]+x_f, y = yadj2+1.6, label = TeX(dat_textcvr$label[1]), size = 8) +
  theme(axis.text = element_text(size = size_axis), 
        axis.text.y = element_text(size = size_axis), 
        axis.title = element_text(size = size_axis), 
        plot.title = element_text(face = "bold", size = 36, hjust = 0.5)) + 
    ylim(-6, 3.5)

# Marine
  x_m = -0.4
  p_m_cv <-  bubble_plot(model4.9_m, mod = "cv_c", group = "spp", ylab = TeX("$lnCVR_{Q_{10_{acclim}}}$"), xlab = "Thermal Coefficient of Variation (CV)", g =FALSE, k = FALSE) + 
    geom_point(data = lnCVRQ10_data_wild_acclim_m, 
                 aes(x = cv_c, y = effect_size, size = 1/sqrt(V)), 
                 fill = col[2], colour = "black", stroke = 0.2, shape = 21) + 
    annotate("text", x = dat_textcvr$x[1]+x_m, y = yadj2, label = TeX(dat_textcvr$p[2]), size = 8) +
    annotate("text", x = dat_textcvr$x[1]+x_m, y = yadj2+0.8, label = TeX(dat_textcvr$label2[2]), size = 8) +
    annotate("text", x = dat_textcvr$x[1]+x_m, y = yadj2+1.6, label = TeX(dat_textcvr$label[2]), size = 8) +
    theme(axis.text = element_text(size = size_axis), 
        axis.text.y = element_text(size = size_axis), 
        axis.title = element_text(size = size_axis), 
        plot.title = element_text(face = "bold", size = 36, hjust = 0.5)) + 
    ylim(-6, 3.5)

@@

x_t = 0
p_t_cv <-  bubble_plot(model4.9_t, mod = "cv_c", group = "spp", ylab = TeX("$lnCVR_{Q_{10_{acclim}}}$"), xlab = "Thermal Coefficient of Variation (CV)", g = FALSE, k = FALSE) + 
  geom_point(data = lnCVRQ10_data_wild_acclim_t, 
               aes(x = cv_c, y = effect_size, size = 1/sqrt(V)), 
               fill = col[3], colour = "black", stroke = 0.2, shape = 21) +
  annotate("text", x = dat_textcvr$x[3]+x_t, y = yadj2, label = TeX(dat_textcvr$p[3]), size = 8) +
  annotate("text", x = dat_textcvr$x[3]+x_t, y = yadj2+0.8, label = TeX(dat_textcvr$label2[3]), size = 8) +
  annotate("text", x = dat_textcvr$x[3]+x_t, y = yadj2+1.6, label = TeX(dat_textcvr$label[3]), size = 8) +
  theme(axis.text = element_text(size = size_axis), 
      axis.text.y = element_text(size = size_axis), 
      axis.title = element_text(size = size_axis), 
      plot.title = element_text(face = "bold", size = 36, hjust = 0.5)) + 
  ylim(-6, 3.5)

@@

  p_cv <- (p_t_rr + p_m_rr + p_f_rr) / (p_t_cv + p_m_cv + p_f_cv) + 
            plot_annotation(tag_levels = "A") & theme(plot.tag = element_text(size = 36))

@@

```{r, global}
#| label: global
#| echo: false
#| include: false

@@

@@

@@

@@

```{r, future_clim}
#| label: future_clim
#| echo: false
#| warning: false
#| include: false

@@

@@

@@

@@

Measurements of acute and acclimation responses from wild ectotherms
were much less common than from captive populations ($N_{species}$ =
`r length(unique(all_pred$spp))`, from
`r length(unique(all_pred$uniques))` wild populations). Globally, there
was a clear bias towards species in the Northern Hemisphere
([@fig-fig6]A-C). Projected changes in physiological rate variance were
highly variable across the globe, with some regions showing a decrease
in physiological rate variance, while others showing an increase
([@fig-fig6]D). However, out of the `r length(unique(all_pred2$match))`
populations, relative variance was predicted to decrease in
`r (sum(decreases$n, na.rm = TRUE) / length(decreases$n[complete.cases(decreases$n)]))*100`%
of the locations.

Using the ERA5 climate model, predictions of current global changes in
physiological rate variance were generally conservative with our model
explaining \~ 50% of the variation in the observed data ($R^2$ =
`r round(mean(r2_2), 2)`, 95% CI: `r round(quantile(r2_2, 0.025)[1], 2)`
to `r round(quantile(r2_2, 0.975)[1], 2)`). Climate change is predicted
to result in a `r abs(lapply(mean_perc_habitat, mean)$f)`% reduction in
relative variance for freshwater systems (95% CI:
`r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$f[1]`
to
`r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$f[2]`%,
$p_{MCMC}$ = `r pmcmc(mean_perc_habitat$f)`), a
`r abs(lapply(mean_perc_habitat, mean)$m)`% reduction in relative
variance for marine systems (95% CI:
`r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$m[1]`
to
`r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$m[2]`%,
$p_{MCMC}$ = `r pmcmc(mean_perc_habitat$m)`), and a
`r abs(lapply(mean_perc_habitat, mean)$t)`% reduction in relative
variance for terrestrial systems (95% CI:
`r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$t[1]`
to
`r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$t[2]`%,
$p_{MCMC}$ = `r p_value(pmcmc(mean_perc_habitat$t))`) under a RCP8.5
climate scenario. All results are taken from "Model 6".

```{r, fig-fig6}
#| label: fig-fig6
#| echo: false
#| warning: false
#| message: false
#| fig-cap: "Model predictions for the expected change in acclimation $lnCVR_{Q_{10}}$ across the globe for terrestrial, marine and freshwater ecthotherms. Predictions consider the uncertainty in random effects (i.e., species, phylogeny, study).  Predicted change in relative physiological rate variance for each population based on current temperatures (average from 2018-2022; A-C) as well as the expected change from current temperatures based on future temperature predictions (average from 2096-2100, D). Future climate predictions are the reduction in relative variance expected under a RCP8.5 climate scenario relative to current climate conditions (% change)."

@@

figvr_lh <-(p3_cvr | p1_cvr | p2_cvr) + plot_annotation(tag_levels = "A") 
ggsave(figcvr_lh, filename = "./output/figures/figSlh.pdf", width = 28, height = 10, device = "pdf")

@@

>>>>

  # First get sample sizes
    N <- lnRRQ10_data %>% group_by(habitat, type) %>% summarise(n=n())

@@

# First get sample sizes
    N <- lnVRQ10_data %>% group_by(habitat, type) %>% summarise(n=n())

@@

# Bring in  historical climate data
  climate <- read.csv("./output/climate_data/ERA5_air_and_SST_timeseries.csv")

@@

############################################################################################
# Model 5.9: Acclimatization lnVRQ10 only. Test if environmental variance explains 
# differently in each habitat
############################################################################################          
  lnVRQ10_data_wild_acclim <- lnVRQ10_data_wild %>%  filter(type == "acclim")

@@

  # Check and prune tree
tree_lnvr_wild_acclim <- tree_checks(data = lnVRQ10_data_wild_acclim, tree = tree_p0.7, dataCol = "species_full", type = "prune")
tree_checks(data = lnVRQ10_data_wild_acclim, tree = tree_lnvr_wild_acclim, dataCol = "species_full", type = "check")

phylo_cor_wild_acclim <- vcv(tree_lnvr_wild_acclim, cor = TRUE)
        A_wild_acclim <- vcv(tree_lnvr_wild_acclim, cor = FALSE) # For bayes, need cov matrix

if(rerun){
  # More realistic model with 
  model5.9  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + cv_c + habitat + habitat:cv_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnVRQ10_data_wild_acclim)

   model5.9_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  max_temp_c + c_accl_time + cv_c + habitat + habitat:cv_c + (~1 | record_num) + (1 | gr(species_full, cov = A_wild_acclim)) + (1 + max_temp_c| spp) + (1|trait), data2 = list(A_wild_acclim=A_wild_acclim), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.99, max_treedepth = 12), chains = 4, thin = 1, iter = 5000, warmup = 1000, cores = 8, data = lnVRQ10_data_wild_acclim)

  # Save models
  saveRDS(model5.9, file = "./output/models/lnVRQ10_5.9")
  saveRDS(model5.9_bayes, file = "./output/models/lnVRQ10_5.9_bayes")

  } else{
        model5.9 <- readRDS("./output/models/lnVRQ10_5.9")
  model5.9_bayes <- readRDS("./output/models/lnVRQ10_5.9_bayes")
}

#################################################
### Bayesian Slopes from Model 5.9
#################################################
 posteriors5.9 <- posterior_samples(model5.9_bayes, "^b")

    slope_f_5.9 <- posteriors5.9[,"b_cv_c"]
    slope_m_5.9 <- posteriors5.9[,"b_cv_c"] + posteriors5.9[,"b_cv_c:habitatm"]
    slope_t_5.9 <- posteriors5.9[,"b_cv_c"] + posteriors5.9[,"b_cv_c:habitatt"]

##################################################
    # Model 5.8, 5.9 and 5.10 - Habitat subset
    # Used for making plots
##################################################
    # Subset data to habitat
    lnVRQ10_data_wild_acclim_t <- lnVRQ10_data_wild_acclim %>% filter(habitat == "t")
    lnVRQ10_data_wild_acclim_f <- lnVRQ10_data_wild_acclim %>% filter(habitat == "f")
    lnVRQ10_data_wild_acclim_m <- lnVRQ10_data_wild_acclim %>% filter(habitat == "m")

if(rerun){
    # Terrestrial Models
    model5.9_t  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + cv_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnVRQ10_data_wild_acclim_t, control=list(iter.max=1000, rel.tol=1e-8))
    saveRDS(model5.9_t, "./output/models/model5.9_t")

@@

    model5.9_f  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + cv_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnVRQ10_data_wild_acclim_f)

    saveRDS(model5.9_f, "./output/models/model5.9_f")

@@

    model5.9_m  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + cv_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnVRQ10_data_wild_acclim_m)
    saveRDS(model5.9_m, "./output/models/model5.9_m")

  } else {
     model5.9_t <- readRDS("./output/models/model5.9_t")
     model5.9_f <- readRDS("./output/models/model5.9_f")
     model5.9_m <- readRDS("./output/models/model5.9_m")
}

@@

####################
# lnVR CV
####################
dat_textvr <- data.frame(label = c(text(slope_f_5.9), 
                                 text(slope_m_5.9), 
                                 text(slope_t_5.9)),
                      label2 = c(text(slope_f_5.9, b = FALSE), 
                                 text(slope_m_5.9,b = FALSE), 
                                 text(slope_t_5.9,b = FALSE)),
                           p = c(p_text(slope_f_5.9), 
                                 p_text(slope_m_5.9), 
                                 p_text(slope_t_5.9)),
                       title = c("Freshwater", "Marine", "Terrestrial"), 
                      condition = c("f", "m", "t"),
                      x = rep(0, 3),
                      y = rep(-4, 3))

# Freshwater
  yadj2 = -6
  x_f = 0.6
  p_f_vr <-  bubble_plot(model5.9_f, mod = "cv_c", group = "spp", ylab = TeX("$lnVR_{Q_{10_{acclim}}}$"), xlab = "Thermal Coefficient of Variation (CV)", g =FALSE, k = FALSE) + 
      geom_point(data = lnVRQ10_data_wild_acclim_f, aes(x = cv_c, y = effect_size, size = 1/sqrt(V)), fill = col[1], colour = "black", stroke = 0.2, shape = 21) + 
    annotate("text", x = dat_textvr$x[1]+x_f, y = yadj2, label = TeX(dat_textvr$p[1]), size = 8) +
    annotate("text", x = dat_textvr$x[1]+x_f, y = yadj2+0.8, label = TeX(dat_textvr$label2[1]), size = 8) +
    annotate("text", x = dat_textvr$x[1]+x_f, y = yadj2+1.6, label = TeX(dat_textvr$label[1]), size = 8) +
  theme(axis.text = element_text(size = size_axis), 
        axis.text.y = element_text(size = size_axis), 
        axis.title = element_text(size = size_axis), 
        plot.title = element_text(face = "bold", size = 36, hjust = 0.5)) + 
    ylim(-6, 3.5)

# Marine
  x_m = -0.4
  p_m_vr <-  bubble_plot(model5.9_m, mod = "cv_c", group = "spp", ylab = TeX("$lnVR_{Q_{10_{acclim}}}$"), xlab = "Thermal Coefficient of Variation (CV)", g =FALSE, k = FALSE) + 
    geom_point(data = lnVRQ10_data_wild_acclim_m, 
                 aes(x = cv_c, y = effect_size, size = 1/sqrt(V)), 
                 fill = col[2], colour = "black", stroke = 0.2, shape = 21) + 
    annotate("text", x = dat_textvr$x[1]+x_m, y = yadj2, label = TeX(dat_textvr$p[2]), size = 8) +
    annotate("text", x = dat_textvr$x[1]+x_m, y = yadj2+0.8, label = TeX(dat_textvr$label2[2]), size = 8) +
    annotate("text", x = dat_textvr$x[1]+x_m, y = yadj2+1.6, label = TeX(dat_textvr$label[2]), size = 8) +
    theme(axis.text = element_text(size = size_axis), 
        axis.text.y = element_text(size = size_axis), 
        axis.title = element_text(size = size_axis), 
        plot.title = element_text(face = "bold", size = 36, hjust = 0.5)) + 
    ylim(-6, 3.5)

@@

x_t = 0
p_t_vr <-  bubble_plot(model5.9_t, mod = "cv_c", group = "spp", ylab = TeX("$lnVR_{Q_{10_{acclim}}}$"), xlab = "Thermal Coefficient of Variation (CV)", g = FALSE, k = FALSE) + 
  geom_point(data = lnVRQ10_data_wild_acclim_t, 
               aes(x = cv_c, y = effect_size, size = 1/sqrt(V)), 
               fill = col[3], colour = "black", stroke = 0.2, shape = 21) +
  annotate("text", x = dat_textvr$x[3]+x_t, y = yadj2, label = TeX(dat_textvr$p[3]), size = 8) +
  annotate("text", x = dat_textvr$x[3]+x_t, y = yadj2+0.8, label = TeX(dat_textvr$label2[3]), size = 8) +
  annotate("text", x = dat_textvr$x[3]+x_t, y = yadj2+1.6, label = TeX(dat_textvr$label[3]), size = 8) +
  theme(axis.text = element_text(size = size_axis), 
      axis.text.y = element_text(size = size_axis), 
      axis.title = element_text(size = size_axis), 
      plot.title = element_text(face = "bold", size = 36, hjust = 0.5)) + 
  ylim(-6, 3.5)

@@

  p_cv <- (p_t_rr + p_m_rr + p_f_rr) / (p_t_vr + p_m_vr + p_f_vr) + 
            plot_annotation(tag_levels = "A") & theme(plot.tag = element_text(size = 36))

@@

```{r, globalvr}
#| label: global
#| echo: false
#| include: false

@@

   model6_bayes <- brms::brm(effect_size | se(sqrt(V)) ~  max_temp_c + c_accl_time + type*habitat + t2(Longitude, Latitude) +
                               (1 | record_num) + 
                               (1 | gr(species_full, cov = A_wild_vr)) + 
                               (1 + max_temp_c | spp) + 
                               (1 | trait) + 
                               (1 | obs), 
                      data2 = list(A_wild_vr=A_wild_vr), family = gaussian(), 
                      prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), 
                      control = list(adapt_delta = 0.99, max_treedepth = 12), 
                      chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, 
                      data = lnVRQ10_data_wild[complete.cases(lnVRQ10_data_wild$c_accl_time),])

  saveRDS(model6_bayes, file = "./output/models/lnVRQ10_6_bayes")

  } else{
  model6_bayes <- readRDS("./output/models/lnVRQ10_6_bayes")
}

# Lets explore the model and how it's predicting the real data
  data_w_vr <- lnVRQ10_data_wild[complete.cases(lnVRQ10_data_wild$c_accl_time),]

# combine with predictions
  data_w_pred_vr <- cbind(data_w_vr, predict(model6_bayes))

# Fitted vs actual data
  ggplot(data = data_w_pred_vr, aes(x = Estimate, y = effect_size))      + 
        geom_point() + geom_smooth(method = "lm")                     + 
        labs(x = "Predicted Effect Size", y = "Observed Effect Size") + 
        theme_classic()                                               + 
        geom_abline(intercept = 0, slope = 1)

@@

#  Spot check of a bunch of studies doesn't show any obvious issues with the data. Note that data is repeated because acute and acclimated rates are calculated. We expect the rows of raw data for acute and acclim to be repeated even though these data are NOT all used to make the calculation.
  sign_change_dat_vr <- data_w_pred_vr  %>% filter(sign(Estimate) != sign(effect_size))
  #View(sign_change_dat)

# Fitted by Lat/long Longitude, Latitude
  ggplot(data = data_w_pred_vr, aes(y = Estimate, x = Latitude)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Latitude", y = "Predicted Effect Size") + theme_classic()

  ggplot(data = data_w_pred_vr, aes(y = Estimate, x = Longitude)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Longitude", y = "Predicted Effect Size") + theme_classic()

# Observed by Lat/long
  ggplot(data = data_w_pred_vr, aes(y = effect_size, x = Latitude)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Latitude", y = "Observed Effect Size") + theme_classic()

  ggplot(data = data_w_pred_vr, aes(y = effect_size, x = Longitude)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Longitude", y = "Observed Effect Size") + theme_classic()

# How much variation does this model explain. Need to do this manually because function doesn't seem to work on the model object. Note that this is just for fixed effects, re_formula = NA ignore the random effects.
  re_post_vr <- posterior_samples(model6_bayes, pars = "^sd_")
    v_fix_vr <- apply(predict(model6_bayes, summary = FALSE, re_formula = NA), 1, function(x) var(x))

# sigma2_v = typical sampling error variance
  sub_data_vr <- lnVRQ10_data_wild[complete.cases(lnVRQ10_data_wild$c_accl_time),]
  sigma2_v_vr <- sum(1 / sub_data_vr$V) * (dim(sub_data_vr)[1] - 1) /
                 (sum(1 / sub_data_vr$V)^2 - sum((1 / sub_data_vr$V)^2))

# Extract the random effect variances
  v_trait_vr <- re_post_vr$sd_trait__Intercept^2
    v_spp_vr <- re_post_vr$sd_spp__Intercept^2
    v_phy_vr <- re_post_vr$sd_species_full__Intercept^2
    v_std_vr <- re_post_vr$sd_record_num__Intercept^2
      v_r_vr <- re_post_vr$sd_obs__Intercept^2

# Lets get the total variance for the random effects
  tot_re_vr <- v_trait_vr + v_spp_vr + v_phy_vr + v_std_vr + v_r_vr

# Now, calculate r2. Note that this isn't perfect and should be under estimated because we are ignoring random slope
         r2 <- (v_fix_vr) / (v_fix_vr + tot_re_vr)

    mean(r2) #  0.43
    quantile(r2, probs = c(0.025, 0.975)); #0.33-0.51

######################################################################
# Model predictions for terrestrial, marine and freshwater ectotherms
######################################################################
# Now we need to make predictions for new data. We first need to create new data. We need to do this for each habitat type because they are all a little different.

@@

  terrest_data_vr <- sub_data_vr  %>% 
                  filter(habitat == "t")  %>% 
                  distinct(spp, species_full, habitat, type, Latitude, Longitude)

  new_data_t <- data.frame(
                            max_temp_c = 0, 
                            c_accl_time = 0,
                            V = 0,
                            trait = 0,
                            obs = 0, 
                            record_num = 0
    )

  # Create new data frame with only unique values for prediction.
  new_data_t2_vr  <- cbind(terrest_data_vr, new_data_t)

  # We'll now make the predictions from the model object using the new data frame
  terre_pred_vr <- predict(model6_bayes, newdata = new_data_t2_vr, 
                        summary = TRUE, re_formula = NA, 
                        robust = FALSE, intervals = "confidence")

  # We'll also keep the whole posterior which we'll use later on. We need to transpose so that the data match. Each numbered columns is one sample of the posterior.
  terre_pred_no_sum_vr <- cbind(new_data_t2_vr, 
                                t(predict(model6_bayes, newdata = new_data_t2_vr, re_formula = NA, summary = FALSE)))

  # Lets merge predictions with the original data
  terrest_data_vr <- cbind(terrest_data_vr, terre_pred_vr)

  # Test manual calculations. We can also set summary = FALSE and then do the calculations ourselves. Note we want robust = FALSE to make the comparisons more similar as otherwise the calculation are different when using summary TRUE/FALSE . These are identical.
  terre_pred2 <- predict(model6_bayes, newdata = new_data_t2_vr, summary = FALSE, re_formula = NA)
  plyr::ldply(apply(terre_pred2, 2, function(x) post_summary(x)))

#############################
## Marine ectotherms  
#############################
  marine_data_vr <- sub_data_vr  %>% filter(habitat == "m") %>% distinct(spp, species_full, habitat, type, Latitude, Longitude)

@@

  # Create new data frame with only unique values for prediction.
  new_data_m_vr  <- cbind(marine_data_vr, new_data_m)

  # We'll now make the predictions from the model object using the new data frame
  marine_pred_vr <- predict(model6_bayes, newdata = new_data_m_vr, summary = TRUE, 
                            re_formula = NA, intervals = "confidence")

  # We'll also keep the whole posterior which we'll use later on. We need to transpose so that the data match. Each numbered columns is one sample of the posterior.
  marine_pred_no_sum_vr <- cbind(new_data_m_vr, t(predict(model6_bayes, newdata = new_data_m_vr, 
                                                          summary = FALSE, re_formula = NA)))

  # Lets merge predictions with the original data
  marine_data_vr <- cbind(marine_data_vr, marine_pred_vr)

#############################
## Freshwater ectotherms
#############################
  fresh_data_vr <- sub_data_vr  %>% 
                  filter(habitat == "f") %>% 
                  distinct(spp, species_full, habitat, type, Latitude, Longitude)

  new_data_f <- data.frame(
                             max_temp_c = 0, 
                            c_accl_time = 0,
                                      V = 0,
                                  trait = 0,
                                    obs = 0, 
                             record_num = 0)

  # Create new data frame with only unique values for prediction.
  new_data_f_vr  <- cbind(fresh_data_vr, new_data_f)

  # We'll now make the predictions from the model object using the new data frame
  fresh_pred_vr <- predict(model6_bayes, newdata = new_data_f_vr, summary = TRUE, 
                           re_formula = NA, intervals = "confidence")

  # We'll also keep the whole posterior which we'll use later on. We need to transpose so that the data match. Each numbered columns is one sample of the posterior.
  fresh_pred_non_sum_vr <- cbind(new_data_f_vr, t(predict(model6_bayes, newdata = new_data_f_vr,
                                                          summary = FALSE, re_formula = NA)))

  # Lets merge predictions with the original data
  fresh_data_vr <- cbind(fresh_data_vr, fresh_pred_vr)

# Now that we have all the predictions we can merge the data together. Note that species itself is not really that useful because each prediction at each latitude and longitude should incorporate the species effect. So we'll remove it. While I retain spp, it's still not totally specific to the species, but left it there so it's clear how these data match with the wild data. Note there are only 134 species because this is subset of the wild data.
         all_pred_vr <- rbind(terrest_data_vr, marine_data_vr, fresh_data_vr) %>% 
                        dplyr::select(-species_full)  %>% 
                        mutate(uniques = as.character(interaction(spp, Latitude, Longitude)))
  all_pred_no_sum_vr <- rbind(terre_pred_no_sum_vr, marine_pred_no_sum_vr, fresh_pred_non_sum_vr) %>%      
                        dplyr::select(-species_full)
           spp_wild  <- length(unique(sub_data_vr$spp))
```

```{r, future_climvr}
#| label: future_clim
#| echo: false
#| warning: false
#| include: false

@@

# Lets clean this data up. We only need the mean temperature here, but this needs to be converted to Celsius as right now it's in Kelvins. We'll now take the last 5 years of the future temperature 2096-2100
  future_temps  <- future_temp  %>% dplyr::select(Species, Latitude, Longitude,  grep("X2096|X2097|X2098|X2099|X2100", colnames(future_temp)))  %>% mutate(mean = apply(.[, grep("X2096|X2097|X2098|X2099|X2100", colnames(.))], 1, function(x) mean(x, na.rm = TRUE)), Celc_future = mean - 273.15,  match = interaction(Species, Latitude, Longitude))

# Now we can merge the climate data sets together so that we have the species, lat and long, current and future climate. Then we can calculate the expected 'increase' in T into the future

@@

# Lets have a global plot of temperature changes to see it spatially

@@

# Now lets merge all_pred_vr with the climate projections
  # Simplify columns 
  climate_data_new <- climate_data  %>% dplyr::select(-Latitude, -Longitude)

  # Merge with model predictions. This is the predicted lnVRQ10 for 10C change AT the specific latitude and longitude. We must assume that there is no plasticity or adaptation at each site to make projections.
  all_pred2_vr <- all_pred_vr %>% mutate(match = interaction(spp, Latitude, Longitude)) %>% 
              left_join(., climate_data_new, by = "match")  %>% 
              mutate(lnVR_future = (Estimate / 10) * T_change,
                      lnVR_F_l95 = (Q2.5 / 10) * T_change,
                      lnVR_F_u95 = (Q97.5 / 10) * T_change, 
                      per_change = (exp(lnVR_future) - 1) * 100,
                      per_change_l95 = (exp(lnVR_F_l95) - 1) * 100,
                      per_change_u95 = (exp(lnVR_F_u95) - 1) * 100,
                     match = as.character(match),
                     uniques = as.character(uniques))

  # First, lets merge the climate data with the predictions, but here predictions are not summarized. We have the entire posterior

    all_pred2_no_sum_vr <-  all_pred_no_sum_vr %>% 
                        mutate(match = interaction(spp, Latitude, Longitude))  %>% 
                        left_join(., climate_data_new, by = "match")

  # Second, lets convert each estimate to lnvr to 1 degree C instead of 10...
   all_pred2_no_sum2_vr <- all_pred2_no_sum_vr %>%  
                        mutate_if(colnames(.) %in% c(1:3200), lncvr_1c) # Note, names lncvr_1c but will work for any effect

  # Third, lets multiply each estimate by the temperature change at that location. I've checked 4 different cells and they are calculated correctly.
    all_pred2_no_sum2_vr[,colnames(all_pred2_no_sum2_vr) %in% c(1:3200)]  <- all_pred2_no_sum2_vr[,colnames(all_pred2_no_sum2_vr) %in% c(1:3200)] * all_pred2_no_sum2_vr$T_change

  # Fourth, we can calculate the percentage change on the posterior
    all_pred2_no_sum2_vr[,paste0("perc", c(1:3200))]  <- (exp(all_pred2_no_sum2_vr[,colnames(all_pred2_no_sum2_vr) %in% c(1:3200)]) - 1)*100

  # Now we can summarise each. Here, what we are doing is taking the mean for each column. Effectively pooling across all the iterations for each dataset. This gives us an estimation of the mean change in variance across all locations/species, but it gives us 3200 point estimates of this mean. We can then use this to calculate the 95% CI's.
    mean_perc <-  apply(all_pred2_no_sum2_vr[,paste0("perc", c(1:3200))], 1, mean, na.rm = TRUE) 
    mean(mean_perc)
    quantile(mean_perc, 0.025)
    quantile(mean_perc, 0.975)
    hist(mean_perc)

  # Now, lets do this but for each habitat, the good old fashioned way. I think that these are so variable because the expected change in the future is quite dramatic, so the variability is high.
  mean_perc_habitat_vr <-  lapply(split(all_pred2_no_sum2_vr,  all_pred2_no_sum2_vr$habitat), function(x) apply(x[,paste0("perc", c(1:3200))], 1, mean, na.rm = TRUE))
  lapply(mean_perc_habitat_vr, mean)
  lapply(mean_perc_habitat_vr, function(x) quantile(x, c(0.025, 0.975)))

# Create a table to summarise everything
tables_future <- all_pred2_vr %>% group_by(habitat) %>% summarise(mean = mean(per_change, na.rm = TRUE), 
                                                                l95 = mean(per_change_l95, na.rm = TRUE), 
                                                                u95 = mean(per_change_u95, na.rm = TRUE))

# Check on how many to increase
  increases <- all_pred2_vr %>% group_by(match)  %>% summarise(n = if_else(any(per_change > 0), 1, 0))

@@

Measurements of acute and acclimation responses from wild ectotherms
were much less common than from captive populations ($N_{species}$ =
`r length(unique(all_pred2_vr$spp))`, from
`r length(unique(all_pred2_vr$uniques))` wild populations). Globally, there
was a clear bias towards species in the Northern Hemisphere
([@fig-fig6]A-C). Projected changes in physiological rate variance were
highly variable across the globe, with some regions showing a decrease
in physiological rate variance, while others showing an increase
([@fig-fig6]D). However, out of the `r length(unique(all_pred2_vr$match))`
populations, variance was predicted to increase to some extent at
`r (sum(increases$n, na.rm = TRUE) / length(increases$n[complete.cases(increases$n)]))*100`%
of the locations.

Using the ERA5 climate model, predictions of current global changes in
physiological rate variance were generally conservative with our model
explaining ~ 43% of the variation in the observed data ($R^2$ =
`r round(mean(r2), 2)`, 95% CI: `r round(quantile(r2, 0.025)[1], 2)`
to `r round(quantile(r2, 0.975)[1], 2)`). Climate change is predicted
to result in a `r abs(lapply(mean_perc_habitat_vr, mean)$f)`% reduction in
relative variance for freshwater systems (95% CI:
`r lapply(mean_perc_habitat_vr, function(x) quantile(x, c(0.025, 0.975)))$f[1]`
to
`r lapply(mean_perc_habitat_vr, function(x) quantile(x, c(0.025, 0.975)))$f[2]`%,
$p_{MCMC}$ = `r pmcmc(mean_perc_habitat_vr$f)`), a
`r abs(lapply(mean_perc_habitat_vr, mean)$m)`% reduction in relative
variance for marine systems (95% CI:
`r lapply(mean_perc_habitat_vr, function(x) quantile(x, c(0.025, 0.975)))$m[1]`
to
`r lapply(mean_perc_habitat_vr, function(x) quantile(x, c(0.025, 0.975)))$m[2]`%,
$p_{MCMC}$ = `r pmcmc(mean_perc_habitat_vr$m)`), and a
`r abs(lapply(mean_perc_habitat_vr, mean)$t)`% reduction in relative
variance for terrestrial systems (95% CI:
`r lapply(mean_perc_habitat_vr, function(x) quantile(x, c(0.025, 0.975)))$t[1]`
to
`r lapply(mean_perc_habitat_vr, function(x) quantile(x, c(0.025, 0.975)))$t[2]`%,
$p_{MCMC}$ = `r p_value(pmcmc(mean_perc_habitat_vr$t))`) under a RCP8.5
climate scenario. All results are taken from "Model 6".

```{r, fig-fig6}
#| label: fig-fig6
#| echo: false
#| warning: false
#| message: false
#| fig-cap: "Model predictions for the expected change in acclimation $lnVR_{Q_{10}}$ across the globe for terrestrial, marine and freshwater ecthotherms. Predictions consider the uncertainty in random effects (i.e., species, phylogeny, study).  Predicted change in relative physiological rate variance for each population based on current temperatures (average from 2018-2022; A-C) as well as the expected change from current temperatures based on future temperature predictions (average from 2096-2100, D). Future climate predictions are the reduction in relative variance expected under a RCP8.5 climate scenario relative to current climate conditions (% change)."

@@

figcvr_lh <-(p3_cvr | p1_cvr | p2_cvr) + plot_annotation(tag_levels = "A") 
ggsave(figcvr_lh, filename = "./output/figures/figSlh.pdf", width = 28, height = 10, device = "pdf")

@@

```{r, globallncvr}
#| label: globallncvr
#| echo: false
#| include: false

```{r, future_climlncvr}
#| label: future_climlncvr
#| echo: false
#| warning: false
#| include: false
